### 2022-07-27 paper/localpsf.tex
adding a scalar descriptor word for the description of the discretization of the pressure field, in order that we do not unnecessarily distinguish this discretization from that of the parameter field
removing an unnecessary comma in between Hgn a qk in equation 6.3 for determining the search direction
removing a symbol for the Glen exponent, as it is only used once, no need to have a symbol for later reference

\begin{subequations}\label{Stokeseqn}
	\begin{align}
	-\nabla \cdot \stress
	&=\bodyforce\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
	\nabla\cdot \velocity
	&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
	\stress\normalvec
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
	\velocity \cdot\normalvec =0 \quad\text{and}\quad \tangentop\left(\stress\normalvec+\exp\left(\basalfriction\right)\velocity\right)
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
	\stress\normalvec+\mu \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
	\end{align}
\end{subequations}
The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field. Here, $\basalfriction$ is the unknown logarithmic
basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$. The quantity $\bodyforce$ is the body force density due to gravity, $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary. 
We employ Glen's flow law \cite{Glen1955}, $\stress= 2\eta
\dot{\strain} -I\pressure$,
which is a constitutive law for ice that relates the stress
tensor, $\stress$, to the strain rate tensor,
$\dot{\strain}= \frac{1}{2}\left(
\nabla\velocity+\nabla\velocity^{\top} \right)$. Here $\eta$ is the effective viscosity and $I$ is the unit matrix. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model. Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.

The pressure, $\pressure$, is discretized with first order continuous Galerkin finite elements defined on a mesh of tetrahedra. The velocity, $\velocity$, is discretized with second order continuous Galerkin finite elements on the same mesh. The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to $\Gamma_b$. The basal boundary, $\Gamma_b$, is a 2D surface embedded in 3D due to the mountain topography. We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate. The parameter $\basalfriction$ is dually viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building our Hessian approximations and defining the regularization. 
%This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.

@@

We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}, using $\basalfriction_0$ as the initial guess. The parameter is updated iteratively according to the formula
$
	\basalfriction_{k+1} = \basalfriction_k + \alpha_k \searchdir_k,
$
where $\searchdir_k$ approximately solves one of the following Newton linear systems,
\begin{equation}
	\label{eq:newton_system}
	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn}, \searchdir_k = - g_k,
\end{equation}
and the step length $\alpha_k \in \mathbb{R}$ is chosen by line search to satisfy the Armijo conditions \cite{ARMIJO}. Here $g_k$ is the gradient of $J$, $H$ is the Hessian, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$. 
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
We use $H_\text{gn}$ for the first five iterations, and $H$ for all subsequent iterations.  The Newton iteration terminates when $||g_k||/||g_0|| < 10^{-8}$. Systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$. PCG only requires applying $H$ (or $H_\text{gn}$) to vectors. Each application of $H$ (or $H_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.

>>>>

\begin{subequations}\label{Stokeseqn}
	\begin{align}
	-\nabla \cdot \stress
	&=\bodyforce\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
	\nabla\cdot \velocity
	&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
	\stress\normalvec
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
	\velocity \cdot\normalvec =0 \quad\text{and}\quad \tangentop\left(\stress\normalvec+\exp\left(\basalfriction\right)\velocity\right)
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
	\stress\normalvec+\mu \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
	\end{align}
\end{subequations}
The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field. Here, $\basalfriction$ is the unknown logarithmic
basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$. The quantity $\bodyforce$ is the body force density due to gravity, $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary. 
We employ Glen's flow law \cite{Glen1955}, $\stress= 2\eta
\dot{\strain} -I\pressure$,
which is a constitutive law for ice that relates the stress
tensor, $\stress$, to the strain rate tensor,
$\dot{\strain}= \frac{1}{2}\left(
\nabla\velocity+\nabla\velocity^{\top} \right)$. Here $\eta$ is the effective viscosity and $I$ is the unit matrix. Glen's exponent has been chosen to be one, which provides for a linear Stokes model. Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.

The pressure, $\pressure$, is discretized with first order scalar continuous Galerkin finite elements defined on a mesh of tetrahedra. The velocity, $\velocity$, is discretized with second order continuous Galerkin finite elements on the same mesh. The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to $\Gamma_b$. The basal boundary, $\Gamma_b$, is a 2D surface embedded in 3D due to the mountain topography. We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate. The parameter $\basalfriction$ is dually viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building our Hessian approximations and defining the regularization. 
%This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.

@@

We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}, using $\basalfriction_0$ as the initial guess. The parameter is updated iteratively according to the formula
$
	\basalfriction_{k+1} = \basalfriction_k + \alpha_k \searchdir_k,
$
where $\searchdir_k$ approximately solves one of the following Newton linear systems,
\begin{equation}
	\label{eq:newton_system}
	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn} \searchdir_k = - g_k,
\end{equation}
and the step length $\alpha_k \in \mathbb{R}$ is chosen by line search to satisfy the Armijo conditions \cite{ARMIJO}. Here $g_k$ is the gradient of $J$, $H$ is the Hessian, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$. 
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
We use $H_\text{gn}$ for the first five iterations, and $H$ for all subsequent iterations.  The Newton iteration terminates when $||g_k||/||g_0|| < 10^{-8}$. Systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$. PCG only requires applying $H$ (or $H_\text{gn}$) to vectors. Each application of $H$ (or $H_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
